---
title: "ggformula Package"
author: "Nicole Hill"
date: "3/2/2020"
output:
    html_document:
    df_print: paged
    theme: yeti
---


```{r include = FALSE}

url <- paste0("https://github.com/DS4PS/Data-Science-Class/blob",
              "/master/DATA/TempeTrafficAccidents.rds?raw=true")

dat <- readRDS(gzcon(url(url)))    
library(ggformula)
```

# Intro

ggformula is an intuative and easy to use graphic function package that is perfect for beginners. For more information on the package, visit: https://github.com/ProjectMOSAIC/ggformula

ggformula will be demonstrated using traffic accident data from Tempe Open Data Portal, available at: https://data.tempe.gov/dataset/high-severity-traffic-crashes-1-08


# Let's Begin! 

Start by loading or installing the package and importing your data set. 

```{r}
library(ggformula) # loads the package

head(dat) # displays the first few rows of your data
```

# Let's Start With A Simple Histogram 

The following code will produce a simple histogram based on the total number of injuries per accident. This histogram will look similar to the bar graphs we created in elementary school. 

```{r}
gf_histogram (~Totalinjuries, data=dat) # this tells R to create a histogram of total injuries based on our data set, "dat"
```

# Let's Add Some Color 

Grey is boring! Let's add some color. Notice that you can change the color of the outline and fill.  
```{r}
gf_histogram (~Totalinjuries, data=dat, color="green", fill="pink")
```

# Let's Give It A Title 

All good graphs have a title so that we know what we are looking at. Add a title using the following code: 

```{r}
gf_histogram (~Totalinjuries, data=dat, color="green", fill="pink") %>% 
  gf_labs(title="Total Injuries Per Car Accident") # note the %>%...this tells R to pass the data from our first piece of code to the next 
```

# More Labeling 

R was able to populate some generic X and Y labels for us, but we can finetune these with the following code: 

```{r}
gf_histogram (~Totalinjuries, data=dat, color="green", fill="pink") %>% 
  gf_labs(title="Total Injuries Per Car Accident", x="Total Injuries") # the y-axis would be relabeled with y="your label here" 
```

# Really Make It Shine 

Our histogram has come a long way! Let's do one last thing to make it look more visualy appealing - we can change the width of the bars to make them wider or slimer. Since our data is only spread across a few bars, let's make them wider for a more polished look. 

```{r}
gf_histogram (~Totalinjuries, data=dat, color="green", fill="pink", binwidth=1) %>% 
  gf_labs(title="Total Injuries Per Car Accident", x="Total Injuries") #binwidth will determine how wide or slim the bars are...the bigger the number, the bigger the bar.
```

